cvx-flats

### 2020-05-20 neurips-20/cvx-flats-statements.tex
added cts-valued stuff

>>>>

\newcommand{\lbar}{\underline{L}} % couldn't do L* while proofreading...
\newcommand{\iden}{\mathrm{iden}}
\newcommand{\Var}{\mathrm{Var}}

\begin{lemma}[\cite{frongillo2018elicitation}]
  \label{lem:elic-complex-bayes-concave}
  Suppose the loss $L$ elicits $\Gamma:\simplex\to\R$.
  Let $\lbar$ be the Bayes risk of $L$.
  Then for any $p,p'\in\simplex$ with $\Gamma(p)\neq\Gamma(p')$, we have $\lbar(\lambda p + (1-\lambda) p') > \lambda \lbar(p) + (1-\lambda) \lbar(p')$ for all $\lambda\in(0,1)$.
\end{lemma}

\begin{definition}
  A property $\Gamma:\simplex\to\R$ is \emph{$d$-identifiable} if its level sets are flats of co-dimension at most $d$ intersected with $\simplex$.
  \raft{Update to be consistent with our definitions / notation / conventions.  I'd vote for using co-dimension since it makes so much of the analysis easier...  Could make more formal too.}
  We write $\iden(\Gamma) = \min\{d \in \N : \Gamma\text{ is $d$-identifiable}\}$.
\end{definition}

\raf{A modification of the argument of \citet[Corollary 7]{frongillo2018elicitation}}

\begin{theorem}
  \label{thm:bayes-risk-lower-bound}
  Let $L$ elicit some identifiable $\Gamma:\simplex\to\reals^d$ with $\iden(\Gamma)=d$.
  If $\lbar$ is non-constant on every non-singleton level set of $\Gamma$, then $\eliccvx(\lbar) \geq d+1$.
\end{theorem}
\begin{proof}
  Suppose for a contradiction that we have some convex-elicitable property $\hat\Gamma:\simplex\to\reals^d$ and link $\psi : \reals^d \to \reals$ such that $\lbar = \psi \circ \hat\Gamma$.
  The condition $\elic_\I(\Gamma)=d$ implies that $\Gamma$ is $d$-identifiable but not $(d-1)$-identifiable, and thus there must be some level set $\Gamma_r$ which is a flat of dimension $n-d-1$.

  The proof of \citet[Theorem 4]{frongillo2018elicitation} argues that $\hat\Gamma$ must refine $\Gamma$, in the sense that every level set of $\hat\Gamma$ is contained in a level set of $\Gamma$; for completeness we give the argument here.
  Suppose for a contradiction that we have $p,p'$ with $\hat\Gamma(p)=\hat\Gamma(p')$ but $\Gamma(p) \neq \Gamma(p')$.
  As $\lbar = \psi \circ \hat\Gamma$, we also have $\lbar(p) = \lbar(p')$.
  Letting $p'' = \tfrac 1 2 p +  \tfrac 1 2 p'$, Lemma~\ref{lem:elic-complex-bayes-concave} would then give us $\lbar(p'') >  \tfrac 1 2 \lbar(p) +  \tfrac 1 2 \lbar(p') = \lbar(p)$.
  By \citet{osband1985providing}, the level sets $\hat\Gamma_{\hat r}$ are convex, giving $\hat\Gamma(p'') = \hat\Gamma(p)$, which would imply $\lbar(p'')=\lbar(p)$, contradicting $\lbar = \psi \circ \hat\Gamma$.
  We conclude $\hat\Gamma$ must refine $\Gamma$, and in particular, we must have $\hat\Gamma_{\hat r} \subseteq \Gamma_r$ for some $\hat r\in\hat\Gamma(\simplex)$.
  % The proof of \cite[Theorem 4]{frongillo2018elicitation} argues that $\hat\Gamma$ must refine $\Gamma$, in the sense that for all $\hat r \in \hat\Gamma(\simplex)$ we have $\hat\Gamma_{\hat r} \subseteq \Gamma_r$ for some $r\in\reals^k$.

  By Theorem~\ref{thm:cvx-flats}, as $\hat\Gamma$ is $d$-convex elicitable $\hat\Gamma_{\hat r}$ is the union of flats of dimension $n-d-1$ interesected with $\simplex$.
  \raft{Should flesh out this next step; this is the key step}
  As $\hat\Gamma_{\hat r} \subseteq \Gamma_r$ and $\Gamma_r$ is a flat of dimension $n-d-1$ interesected with $\simplex$, we conclude $\hat\Gamma_{\hat r} = \Gamma_r$.
  Since by assumption $\lbar$ is non-constant on $\Gamma_r = \hat\Gamma_{\hat r}$, we have distributions $p,p' \in\hat\Gamma_{\hat r}$ with $\lbar(p)\neq\lbar(p')$, which again contradicts $\lbar = \psi \circ \hat\Gamma$.
\end{proof}

\subsection{Examples \raf{all informal for now}}

Both these results exactly match what Ian and I showed for the class of identifiable properties, which is super interesting...
\begin{itemize}
\item Variance: $\iden(\E Y) = 1$, and over $\simplex$ every non-singleton level set of $\E Y$ has different variances, so the theorem gives $\eliccvx(\Var) \geq 2$.  We then have the matching upper bound since $L(r,y) = (r_1-y)^2 + (r_2-y^2)^2$ is convex in $r$, and we can link to $\Var$ from there.
\item Entropy and norms: In fact, \emph{any} strictly convex/concave function $G$ of $p$ will do here.  You can write it as the Bayes risk (or negative Bayes risk) of a strictly proper scoring rule, which elicits the full distribution $\Gamma(p) = p$.  Since $\iden(\Gamma) = n-1$ since you don't actually need the $n$th coordinate, you have $\eliccvx(G) = n-1$; the upper bound comes from the fact that you can elicit the full distribution with a convex loss and just compute $G$.
\end{itemize}
